---
title: "Visualizations for Distribution of Different Indicators"
output: 
    html_document:
        toc: TRUE
        toc_float: TRUE
        lib_dir: "libs"
        self_contained: FALSE
---

```{r setup, message=FALSE, echo=FALSE, warning=FALSE}
library(data.table)
library(tidytext)
library(tidyverse)
library(dplyr)
library(ggplot2)
library(tidyr)
library(car)
library(readr)
library(lubridate)
library(ggpubr)
library(splines)
library(mgcv)
library(kableExtra)
library(httr)
library(jsonlite)
library(purrr)
library(ggcorrplot)
library(scales)
library(tm)
library(SnowballC)
library(wordcloud2)
library(tibble)
library(caret)
library(randomForest)
library(slam)
library(plotly)
```

It is important to note that, the economic indicators such as GDP tend to have highly skewed distributions, meaning that a few extremely large values dominate the dataset. This creates problems when analyzing relationships between variables, as a small number of high-GDP countries can distort trends, making it difficult to see meaningful patterns. Therefore, I log transformed the economic indicators for better visualization.

Here are the distributions of the three economic indicators across all countries from 1960 to 2023.


```{r, echo=FALSE, warning=FALSE, message=FALSE}
life_expectancy_data <- read.csv("life_expectancy_data.csv")
mortality_data <- read.csv("mortality_data.csv")
immunization_data <- read.csv("immunization_data.csv")
water_access_data <- read.csv("water_access_data.csv")
sanitation_access_data <- read.csv("sanitation_access_data.csv")
gdp_data <- read.csv("gdp_data.csv")
unemployment_data <- read.csv("unemployment_data.csv")
inflation_data <- read.csv("inflation_data.csv")

merged_data <- life_expectancy_data %>%
  full_join(mortality_data, by = c("country_code", "country_name", "year")) %>%
  full_join(immunization_data, by = c("country_code", "country_name", "year")) %>%
  full_join(water_access_data, by = c("country_code", "country_name", "year")) %>%
  full_join(sanitation_access_data, by = c("country_code", "country_name", "year")) %>%
  full_join(gdp_data, by = c("country_code", "country_name", "year")) %>%
  full_join(unemployment_data, by = c("country_code", "country_name", "year")) %>%
  full_join(inflation_data, by = c("country_code", "country_name", "year"))

merged_data <- merged_data %>%
  select(
    country_code = country_code,
    country_name = country_name,
    year = year,
    life_expectancy = indicator_value.x,  
    mortality_rate = indicator_value.y,   
    immunization_rate = indicator_value.x.x,  
    water_access = indicator_value.y.y, 
    sanitation_access = indicator_value.x.x.x,  
    gdp = indicator_value.y.y.y,    
    unemployment_rate = indicator_value.x.x.x.x,  
    inflation_rate = indicator_value.y.y.y.y   
  )

```



```{r interactive-histograms, fig.width=7, fig.height=5,echo = FALSE, message=FALSE, warning=FALSE}
merged_data$log_gdp <- log(merged_data$gdp)
merged_data$log_inflation_rate <- log(merged_data$inflation_rate + 1)
merged_data$log_unemployment_rate <- log(merged_data$unemployment_rate + 1)

p1 <- ggplot(merged_data, aes(x = log_gdp)) + 
  geom_histogram(binwidth = 0.2, fill = "purple", color = "black", alpha = 0.7) +
  labs(
    x = "Log of GDP (in US$)", 
    y = "Frequency", 
    caption = "Figure 1: Log-Transformed GDP Distribution (in US$)"
  ) +
  theme_minimal() +
  theme(
    plot.caption = element_text(hjust = 0.5, size = 12, face = "bold"), 
    axis.title = element_text(size = 12)
  )

p2 <- ggplot(merged_data, aes(x = log_unemployment_rate)) + 
  geom_histogram(binwidth = 0.1, fill = "gold", color = "black", alpha = 0.7) +
  labs(
    x = "Log of Unemployment Rate", 
    y = "Frequency", 
    caption = "Figure 2: Log-Transformed Unemployment Rate Distribution"
  ) +
  theme_minimal() +
  theme(
    plot.caption = element_text(hjust = 0.5, size = 12, face = "bold"), 
    axis.title = element_text(size = 12)
  )

p3 <- ggplot(merged_data, aes(x = log_inflation_rate)) + 
  geom_histogram(binwidth = 0.1, fill = "salmon", color = "black", alpha = 0.7) +
  labs(
    x = "Log of Inflation Rate", 
    y = "Frequency", 
    caption = "Figure 3: Log-Transformed Inflation Rate Distribution"
  ) +
  theme_minimal() +
  theme(
    plot.caption = element_text(hjust = 0.5, size = 12, face = "bold"), 
    axis.title = element_text(size = 12)
  )

p4 <- ggplot(merged_data, aes(x = life_expectancy)) + 
  geom_histogram(binwidth = 1, fill = "steelblue", color = "black", alpha = 0.7) +
  labs(x = "Life Expectancy (years)", y = "Frequency", caption = "Figure 4: Life Expectancy Distribution") +
  theme_minimal() +
  theme(plot.caption = element_text(hjust = 0.5, size = 12, face = "bold"), axis.title = element_text(size = 12))

p5 <- ggplot(merged_data, aes(x = mortality_rate)) + 
  geom_histogram(binwidth = 5, fill = "tomato", color = "black", alpha = 0.7) +
  labs(x = "Mortality Rate (per 1,000)", y = "Frequency", caption = "Figure 5: Mortality Rate Distribution") +
  theme_minimal() +
  theme(plot.caption = element_text(hjust = 0.5, size = 12, face = "bold"), axis.title = element_text(size = 12))

p6 <- ggplot(merged_data, aes(x = immunization_rate)) + 
  geom_histogram(binwidth = 2, fill = "forestgreen", color = "black", alpha = 0.7) +
  labs(x = "Immunization Rate (%)", y = "Frequency", caption = "Figure 6: Immunization Rate Distribution") +
  theme_minimal() +
  theme(plot.caption = element_text(hjust = 0.5, size = 12, face = "bold"), axis.title = element_text(size = 12))

p7 <- ggplot(merged_data, aes(x = water_access)) + 
  geom_histogram(binwidth = 5, fill = "darkcyan", color = "black", alpha = 0.7) +
  labs(x = "Water Access (%)", y = "Frequency", caption = "Figure 7: Water Access Distribution") +
  theme_minimal() +
  theme(plot.caption = element_text(hjust = 0.5, size = 12, face = "bold"), axis.title = element_text(size = 12))

p8 <- ggplot(merged_data, aes(x = sanitation_access)) + 
  geom_histogram(binwidth = 5, fill = "orange", color = "black", alpha = 0.7) +
  labs(x = "Sanitation Access (%)", y = "Frequency", caption = "Figure 8: Sanitation Access Distribution") +
  theme_minimal() +
  theme(plot.caption = element_text(hjust = 0.5, size = 12, face = "bold"), axis.title = element_text(size = 12))


ggplotly(p1)
```

- **Figure 1**: This histogram illustrates the distribution of the log-transformed GDP across countries. The transformation helps normalize skewness from extremely large GDP values, making the distribution more symmetrical and suitable for modeling.


```{r fig.width=7, fig.height=5,echo = FALSE, message=FALSE, warning=FALSE}
ggplotly(p2)
```

- **Figure 2**: The unemployment rate distribution, after log transformation, shows a clearer view of variability across countries. Most countries fall within a relatively narrow range, though some outliers remain.


```{r fig.width=7, fig.height=5,echo = FALSE, message=FALSE, warning=FALSE}
ggplotly(p3)
```

- **Figure 3**: This plot shows the log-transformed inflation rate. The distribution reveals that while most countries have moderate inflation, a few experience significantly higher rates, even after transformation.


Here are the distributions of the five health indicators across all countries from 1960 to 2023.


```{r fig.width=7, fig.height=5,echo = FALSE, message=FALSE, warning=FALSE}
ggplotly(p4)
```

- **Figure 4**: This histogram shows the distribution of life expectancy across countries. Most countries have a life expectancy between 60 and 80 years, indicating relatively high health standards globally.


```{r fig.width=7, fig.height=5,echo = FALSE, message=FALSE, warning=FALSE}
ggplotly(p5)
```

- **Figure 5**: The mortality rate (under age 5) varies significantly. While many countries have low rates, a few exhibit extremely high mortality, reflecting healthcare disparities.


```{r fig.width=7, fig.height=5,echo = FALSE, message=FALSE, warning=FALSE}
ggplotly(p6)
```

- **Figure 6**: Immunization rates cluster around 80–100%, suggesting that many countries have well-established vaccination programs. However, the presence of lower values points to regions with less access.


```{r fig.width=7, fig.height=5,echo = FALSE, message=FALSE, warning=FALSE}
ggplotly(p7)
```

- **Figure 7**: Access to safe water is generally high, but a notable number of countries still fall below universal coverage, indicating room for infrastructure improvement.


```{r fig.width=7, fig.height=5,echo = FALSE, message=FALSE, warning=FALSE}
ggplotly(p8)
```

- **Figure 8**: Sanitation access shows a wider spread, with more countries lacking full coverage. This suggests that sanitation remains a key development challenge.
